Characterizing Optimizations to Memory Access Patterns using Architecture-Independent Program Features

### 2020-03-13 paper/paper.tex
Working on improvements for camera ready

	\subsection{Alignment of Memory Allocation}

@@

	\begin{lstlisting}
		__local float aTile[TILE_DIM][TILE_DIM] __attribute__ ((aligned(4096)));
		__local float bTile[TILE_DIM][TILE_DIM] __attribute__ ((aligned(4096)));
	\end{lstlisting}

	All the examples above are an incomplete listing of optimization strategies developers can apply to code targeted at NVIDIA GPUs.
	In section~\ref{method}, we will compare the collected AIWC metrics with the expected effect of each optimization to examine how effectively our methodology uncovers underlying bottlenecks and guides optimization efforts.

@@

	Our approach is an architecture-independent analysis of memory access patterns to provide metrics correlating to similar performance critical memory access optimizations as CuMAPz.
	We aim to further the state of the art by providing a framework to guide developers in optimizing OpenCL codes for any given target architecture.
	To the best our knowledge, none of the previous works presents a set of performance metrics that accurately characterize memory access patterns of parallel applications independent of the target architecture.

@@

	\begin{figure*}[ht!]
		\centering
		\subfloat{
			\centering
			\begin{tikzpicture}
			\begin{axis}[symbolic x coords={simple,coalA,coalAB,coalABT,alignABT},
			xtick={simple,coalA,coalAB,coalABT,alignABT},
			xticklabel style={text height=2ex},
			ymin=0,
			ylabel={Execution time (ms)}, xlabel={Kernel},
			ylabel style={at={(axis description cs:0.1,.5)},anchor=south},
			legend entries={$N=80\ \ \ \ $,$N=256\ \ $},
			legend style={at={(0.6,0.45)},anchor=west}
			]

			\addplot coordinates {%[mark=*] coordinates {
				(simple,.031)
				(coalA,.019)
				(coalAB,.020)
				(coalABT,.018)
				(alignABT,.018)
			};
			\addplot coordinates {%[mark=star] coordinates {
				(simple,.325)
				(coalA,.075)
				(coalAB,.081)
				(coalABT,.055)
				(alignABT,.055)
			};	
			\end{axis}
			\end{tikzpicture} 
		}
		\subfloat{
			\begin{tikzpicture}
			\begin{axis}[symbolic x coords={simple,coalA,coalAB,coalABT,alignABT},
			xtick={simple,coalA,coalAB,coalABT,alignABT},
			xticklabel style={text height=2ex},
			ymin=0,
			ylabel={Execution time (ms)}, xlabel={Kernel},
			ylabel style={at={(axis description cs:0.1,.5)},anchor=south},
			legend entries={$N=1408$,$N=2048$},
			legend style={at={(0.6,0.45)},anchor=west},
			scaled ticks=false
			]

			\addplot [mark=*, color=brown] coordinates {%[mark=triangle*] coordinates {
				(simple,46.920)
				(coalA,8.783)
				(coalAB,9.238)
				(coalABT,5.307)
				(alignABT,5.272)
			};
			\addplot [mark=star] coordinates {%[mark=square] coordinates {
				(simple,130.682)
				(coalA,28.602)
				(coalAB,27.939)
				(coalABT,16.179)
				(alignABT,16.071)
			};	
			\end{axis}
			\end{tikzpicture} 
		}
		\caption{Execution time of matrix multiplication kernels (NVIDIA Tesla P100)}
		\label{fig: matmul runtime}
	\end{figure*}

@@

	\begin{table*}[h!]
	\centering
	\begin{tabular}[t]{lrrrrr}
		\toprule
		& simple & coalescedA & coalescedAB & coalescedABT & alignedABT \\ \midrule 
		Total memory footprint        & 196608 & 196608     & 196608      & 196608       & 196608     \\ %\hline
		90\% Memory Footprint         & 118196 & 56176      & 489         & 489          & 489        \\ %\hline
		Global MAE                    & 17.02  & 13.18      & 9.78        & 9.78         & 9.78       \\ %\hline
		LMAE \#bits=3                 & 16.02  & 12.18      & 8.78        & 8.78         & 8.78       \\ %\hline
		LMAE \#bits=10                & 9.02   & 5.18       & 1.78        & 1.78         & 1.78       \\ %\hline
		Relative Local Memory Usage   & 0      & 0.50       & 0.94       &  0.94         & 0.94      \\ \bottomrule
	\end{tabular}
	\caption{Selection of AIWC\cite{beauaiwc} metrics for 256 $\times$ 256 matrix multiplication}
	\label{matmul aiwc table}
	\end{table*}

@@

	AIWC's metrics strongly reward optimizations that tend to localize memory accesses. Local memory buffers are typically smaller than global memory arrays when programming for GPUs due to hardware limitations on sizes of shared memory \cite{cudamanual}. However, the metrics currently measured by AIWC do not have a direct causal relation to code patterns that optimize memory accesses on GPUs. The proposed relative local memory usage metric is the first to correspond to a recommended optimization strategy of using fast on-chip shared memory. Further, we find that all current AIWC metrics do not measure any sizeable difference between \texttt{coalescedAB}, \texttt{coalescedABT} and \texttt{alignedABT} codes.
	We address this by proposing another new metric for locality of memory accesses in the following section.

@@

		\begin{figure*}[h!]
		\centering
		\begin{tikzpicture}
		\begin{axis}[symbolic x coords={0,1,2,3,4,5,6,7,8,9,10},
		width=0.7\textwidth,
		height=0.5\textwidth,
		xtick={0,1,2,3,4,5,6,7,8,9,10},
		xticklabel style={text height=2ex},
		ymin=0,
		ylabel={$n$-bits dropped Parallel Spatial Locality}, xlabel={\# of bits skipped},
		ylabel style={at={(axis description cs:0.05,.5)},anchor=south},
		legend entries={\texttt{simple},\texttt{coalescedA},
			\texttt{coalescedAB}, \texttt{coalescedABT}, \texttt{alignedABT}},
		legend style={at={(0.55,0.75)},anchor=west}
		]

@@

		\addplot [mark=*, color=red] coordinates {%[mark=square] coordinates {
			(0,4.145)
			(1,4.145)
			(2,4.145)
			(3,3.2)
			(4,2.254)
			(5,1.308)
			(6,0.3621)
			(7,0.318)
			(8,0.2738)
			(9,0.2296)
			(10,0.1854)
		};

		\addplot coordinates {%[mark=o] coordinates {
			(0,4.145)
			(1,4.145)
			(2,4.145)
			(3,3.2)
			(4,2.254)
			(5,1.308)
			(6,0.3621)
			(7,0.318)
			(8,0.2738)
			(9,0.2296)	
			(10,0.1854)
		};
		\end{axis}
		\end{tikzpicture}
		\caption{Parallel spatial locality metric obtained from AIWC for matrix multiply kernels for 256 $\times$ 256 matrix multiplication}
		\label{fig: matmul parallel spatial locality}
	\end{figure*}

@@

	Figure~\ref{fig: matmul parallel spatial locality} shows the proposed parallel spatial locality metric as measured by AIWC for each of the matrix multiplication kernels from section~\ref{motivating example}.
	We observe that the \texttt{coalescedABT} and \texttt{alignedABT} kernels have the steepest reductions in entropy as the number of bits skipped is increased, which correlates to better locality of parallel memory accesses.
	It is expected for these kernels to exhibit better parallel spatial locality than \texttt{simple}, as \texttt{coalescedABT} and \texttt{alignedABT} make use of local memory, where accesses tend to be localized simply due to the small size of shared on-chip memory typically available on GPUs.
	Further we find that the proposed metric successfully distinguishes between the \texttt{coalescedAB} and \texttt{coalescedABT} kernels.
	It accurately depicts a steeper reduction for the more optimized \texttt{coalescedABT} kernel, where a larger proportion of parallel memory accesses make better use of the memory bank structure of GPU shared memory than all previous kernels.
	This is a significant improvement over the state-of-the-art AIWC metrics in characterizing how codes localize simultaneous memory accesses to better use the hardware provided.

@@

		\addplot coordinates {%[mark=*] coordinates {
			(0, 0.1139)
			(1, 0.1139)
			(2, 0.1139)
			(3, 0.0989)
			(4, 0.08402)
			(5, 0.06914)
			(6, 0.05431)
			(7, 0.03938)
			(8, 0.02442)
			(9, 0.0124)
			(10, 0.00038)
		};

		\addplot coordinates {%[mark=star] coordinates {
			(0, 2.175)
			(1, 2.175)
			(2, 2.175)
			(3, 2.175)
			(4, 2.175)
			(5, 2.175)
			(6, 2.175)
			(7, 2.175)
			(8, 2.175)
			(9, 2.175)
			(10, 2.175)
		};

		\addplot coordinates {%[mark=*] coordinates {
			(0, 7.176)
			(1, 7.176)
			(2, 7.176)
			(3, 6.301)
			(4, 5.418)
			(5, 4.533)
			(6, 3.648)
			(7, 3.619)
			(8, 3.605)
			(9, 3.598)
			(10, 3.594)
		};

		\addplot coordinates {%[mark=star] coordinates {
			(0, 7.097)
			(1, 7.097)
			(2, 7.097)
			(3, 6.223)
			(4, 5.344)
			(5, 4.464)
			(6, 3.583)
			(7, 3.567)
			(8, 3.559)
			(9, 3.554)
			(10, 3.553)
		};

		\addplot coordinates {%[mark=triangle*] coordinates {
			(0, 9.605)
			(1, 9.605)
			(2, 9.605)
			(3, 9.478)
			(4, 9.293)
			(5, 9.049)
			(6, 8.761)
			(7, 8.447)
			(8, 8.12)
			(9,7.233)
			(10,6.287)
		};

		\addplot coordinates { %[mark=square] coordinates {
			(0,4.933)	
			(1,4.933)
			(2,4.933)
			(3,4.89)
			(4,4.795)
			(5,4.652)
			(6,4.456)
			(7,4.215)
			(8,3.93)
			(9,3.608)
			(10,3.253)
		};

		\addplot coordinates { %[mark=square] coordinates {
			(0,4.691)
			(1,4.691)
			(2,4.691)
			(3,4.276)
			(4,3.760)
			(5,3.171)
			(6,2.552)
			(7,1.922)
			(8,1.286)
			(9,0.648)
			(10,0.00776)
		};

		\addplot coordinates { %[mark=o] coordinates {
			(0,1.548)
			(1,1.548)
			(2,1.548)
			(3,1.305)
			(4,1.092)
			(5,0.9385)
			(6,0.8796)
			(7,0.8796)
			(8,0.8796)
			(9,0.8796)
			(10,0.8796)
		};

	    \addplot coordinates { %[mark=star] coordinates {
	    	(0,3.803)
	    	(1,3.803)
	    	(2,3.803)
	    	(3,3.473)
	    	(4,3.144)
	    	(5,2.814)
	    	(6,2.485)
	    	(7,2.485)
	    	(8,2.485)
	    	(9,2.485)
	    	(10,2.485)
    	};

   		\addplot coordinates { %[mark=square*] coordinates {
   			(0,4.101)
   			(1,4.101)
   			(2,4.101)
   			(3,3.589)
   			(4,3.077)
   			(5,2.565)
   			(6,2.053)
   			(7,2.053)
   			(8,2.053)
   			(9,2.053)
   			(10,2.053)
   		};

@@

	%The use of creates imbalanced workload and divergent control flow within work-items in each kernel iteration. While the presence of conditional statements cause some thread divergence in the LUD diagonal kernel, the use of barriers following conditional statements synchronizes the threads. LUD diagonal and perimeter kernels , while LUD.
	%TODO: Divergence in the kernels. Make a diagram/code simplification for LUD.

@@

@@

	To the best of our knowledge, this work is the first to propose the use of architecture-independent metrics of parallel memory access to guide hardware-specific optimization. 
	We implemented and evaluated a new feature, relative local memory usage, to help characterize memory-based GPU code optimizations.
	We also implemented a new parallel spatial locality metric to capture the idea of closeness of memory accesses made by parallel OpenCL workloads.
	We ran the enhanced AIWC tool on matrix multiply kernels and selected OpenDwarfs benchmarks, presenting results and analysis to validate our methodology. 
	%Our work is the first to measure memory access patterns of parallel codes in an architecture-independent, easily interpreted manner. %Josh, Beau: is this too bold a claim? If unnecessary, please remove.
	%TODO UPDATE IN FUTURE

@@

	While our work in this paper intended to guide optimization efforts, another potential use case for AIWC is providing architecture-independent performance predictions for OpenCL kernels by generating machine learning models based on AIWC metrics \cite{beauperformanceprediction}.  Future work would modify the presented metrics and develop new memory metrics with the specific intent of being fed to machine learning models to predict kernel performance on arbitrary and novel architectures. %FIXME: I want to say that we have been focusing on _guiding optimization_. We can do more work in making the AIWC metrics better for _performance prediction_. Is this coming across okay in this paragraph?

>>>>

	\begin{comment}
	\subsection{Alignment of Memory Allocation}

@@

	\begin{lstlisting}
		__local float aTile[TILE_DIM][TILE_DIM] __attribute__ ((aligned(4096)));
		__local float bTile[TILE_DIM][TILE_DIM] __attribute__ ((aligned(4096)));
	\end{lstlisting}
	\end{comment}
	All the examples above are an incomplete listing of optimization strategies developers can apply to code targeted at NVIDIA GPUs.
	In section~\ref{method}, we will compare the collected AIWC metrics with the expected effect of each optimization to examine how effectively our methodology uncovers underlying bottlenecks and guides optimization efforts.

@@

	%TODO: Are we making our work sound too much like CuMAPz? This is of course not true and we should fix Reviewer 2's confusion as to what we are doing differently.

	Our approach is an architecture-independent analysis of memory access patterns to provide metrics correlating to similar performance critical memory access optimizations as CuMAPz.
	We aim to further the state of the art by providing a framework to guide developers in optimizing OpenCL codes for any given target architecture.
	To the best of our knowledge, none of the previous works presents a set of performance metrics that accurately characterize memory access patterns of parallel applications independent of the target architecture.

@@

	\begin{figure*}[ht!]
		\centering
		\subfloat{
			\centering
			\begin{tikzpicture}
			\begin{axis}[symbolic x coords={simple,coalA,coalAB,coalABT,alignABT},
			xtick={simple,coalA,coalAB,coalABT},%alignABT},
			xticklabel style={text height=2ex},
			ymin=0,
			ylabel={Execution time (ms)}, xlabel={Kernel},
			ylabel style={at={(axis description cs:0.1,.5)},anchor=south},
			legend entries={$N=80\ \ \ \ $,$N=256\ \ $},
			legend style={at={(0.6,0.45)},anchor=west}
			]

			\addplot coordinates {%[mark=*] coordinates {
				(simple,.031)
				(coalA,.019)
				(coalAB,.020)
				(coalABT,.018)
				%(alignABT,.018)
			};
			\addplot coordinates {%[mark=star] coordinates {
				(simple,.325)
				(coalA,.075)
				(coalAB,.081)
				(coalABT,.055)
				%(alignABT,.055)
			};	
			\end{axis}
			\end{tikzpicture} 
		}
		\subfloat{
			\begin{tikzpicture}
			\begin{axis}[symbolic x coords={simple,coalA,coalAB,coalABT},%alignABT},
			xtick={simple,coalA,coalAB,coalABT},%alignABT},
			xticklabel style={text height=2ex},
			ymin=0,
			ylabel={Execution time (ms)}, xlabel={Kernel},
			ylabel style={at={(axis description cs:0.1,.5)},anchor=south},
			legend entries={$N=1408$,$N=2048$},
			legend style={at={(0.6,0.45)},anchor=west},
			scaled ticks=false
			]

			\addplot [mark=*, color=brown] coordinates {%[mark=triangle*] coordinates {
				(simple,46.920)
				(coalA,8.783)
				(coalAB,9.238)
				(coalABT,5.307)
				%(alignABT,5.272)
			};
			\addplot [mark=star] coordinates {%[mark=square] coordinates {
				(simple,130.682)
				(coalA,28.602)
				(coalAB,27.939)
				(coalABT,16.179)
				%(alignABT,16.071)
			};	
			\end{axis}
			\end{tikzpicture} 
		}
		\caption{Execution time of matrix multiplication kernels (NVIDIA Tesla P100)}
		\label{fig: matmul runtime}
	\end{figure*}

@@

	\begin{table*}[h!]
	\centering
	\begin{tabular}[t]{lrrrrr}
		\toprule
		& simple & coalescedA & coalescedAB & coalescedABT \\ \midrule %& alignedABT \\ \midrule 
		Total memory footprint        & 196608 & 196608     & 196608      & 196608      \\ %\hline
		90\% Memory Footprint         & 118196 & 56176      & 489         & 489         \\ %\hline
		Global MAE                    & 17.02  & 13.18      & 9.78        & 9.78        \\ %\hline
		LMAE \#bits=3                 & 16.02  & 12.18      & 8.78        & 8.78        \\ %\hline
		LMAE \#bits=10                & 9.02   & 5.18       & 1.78        & 1.78        \\ %\hline
		Relative Local Memory Usage   & 0      & 0.50       & 0.94       &  0.94        \\ \bottomrule
	\end{tabular}
	\caption{Selection of AIWC\cite{beauaiwc} metrics for 256 $\times$ 256 matrix multiplication}
	\label{matmul aiwc table}
	\end{table*}

@@

	AIWC's metrics strongly reward optimizations that tend to localize memory accesses. 
	Local memory buffers are typically smaller than global memory arrays when programming for GPUs due to hardware limitations on sizes of shared memory \cite{cudamanual}. 
	However, the metrics currently measured by AIWC do not have a direct causal relation to code patterns that optimize memory accesses on GPUs. 
	The proposed relative local memory usage metric is the first to correspond to a recommended optimization strategy of using fast on-chip shared memory. 
	Further, we find that all current AIWC metrics do not measure any sizeable difference between \texttt{coalescedAB} and \texttt{coalescedABT} codes.%and \texttt{alignedABT} codes.
	We address this by proposing another new metric for locality of memory accesses in the following section.

@@

		\begin{figure*}[h!]
		\centering
		\begin{tikzpicture}
		\begin{axis}[symbolic x coords={0,1,2,3,4,5,6,7,8,9,10},
		width=0.7\textwidth,
		height=0.5\textwidth,
		xtick={0,1,2,3,4,5,6,7,8,9,10},
		xticklabel style={text height=2ex},
		ymin=0,
		ylabel={$n$-bits dropped Parallel Spatial Locality}, xlabel={\# of bits skipped},
		ylabel style={at={(axis description cs:0.05,.5)},anchor=south},
		legend entries={\texttt{simple},\texttt{coalescedA},
			\texttt{coalescedAB}, \texttt{coalescedABT}},%\texttt{alignedABT}},
		legend style={at={(0.55,0.75)},anchor=west}
		]

@@

		\addplot coordinates {%[mark=square] coordinates {
			(0,4.145)
			(1,4.145)
			(2,4.145)
			(3,3.2)
			(4,2.254)
			(5,1.308)
			(6,0.3621)
			(7,0.318)
			(8,0.2738)
			(9,0.2296)
			(10,0.1854)
		};

		\end{axis}
		\end{tikzpicture}
		\caption{Parallel spatial locality metric obtained from AIWC for matrix multiply kernels for 256 $\times$ 256 matrix multiplication}
		\label{fig: matmul parallel spatial locality}
	\end{figure*}

@@

	Figure~\ref{fig: matmul parallel spatial locality} shows the proposed parallel spatial locality metric as measured by AIWC for each of the matrix multiplication kernels from section~\ref{motivating example}.
	We observe that the \texttt{coalescedABT} kernel has the steepest reductions in entropy as the number of bits skipped is increased, which correlates to better locality of parallel memory accesses.
	It is expected for this kernel to exhibit better parallel spatial locality than \texttt{simple}, as \texttt{coalescedABT} makes use of local memory, where accesses tend to be localized simply due to the small size of shared on-chip memory typically available on GPUs.
	Further we find that the proposed metric successfully distinguishes between the \texttt{coalescedAB} and \texttt{coalescedABT} kernels.
	It accurately depicts a steeper reduction for the more optimized \texttt{coalescedABT} kernel, where a larger proportion of parallel memory accesses make better use of the memory bank structure of GPU shared memory than all previous kernels.
	This is a significant improvement over the state-of-the-art AIWC metrics in characterizing how codes localize simultaneous memory accesses to better use the hardware provided.

@@

		\addplot coordinates {%[mark=*] coordinates {
			(0, 0.08604)
			(1, 0.08604)
			(2, 0.08604)
			(3, 0.07516)
			(4, 0.06431)
			(5, 0.05346)
			(6, 0.04262)
			(7, 0.03176)
			(8, 0.021250)
			(9, 0.01073)
			(10, 0.000218)
		};

		\addplot coordinates {%[mark=star] coordinates {
			(0, .5315)
			(1, .5315)
			(2, .5315)
			(3, .5315)
			(4, .5315)
			(5, .5315)
			(6, .5315)
			(7, .5315)
			(8, .5315)
			(9, .5315)
			(10, .5315)
		};

		\addplot coordinates {%[mark=star] coordinates {
			(0, .8963)
			(1, .8963)
			(2, .8963)
			(3, .7871)
			(4, .6768)
			(5, .5663)
			(6, .4557)
			(7, .4521)
			(8, .4503)
			(9, .4494)
			(10, .4489)
		};

		\addplot coordinates {%[mark=*] coordinates {
			(0, .8866)
			(1, .8866)
			(2, .8866)
			(3, .7773)
			(4, .6675)
			(5, .5576)
			(6, .4476)
			(7, .4456)
			(8, .4446)
			(9, .4441)
			(10, .4439)
		};

		\addplot coordinates {%[mark=triangle*] coordinates {
			(0, .9605)
			(1, .9605)
			(2, .9605)
			(3, .9478)
			(4, .9293)
			(5, .9049)
			(6, .8761)
			(7, .8447)
			(8, .812)
			(9,.7233)
			(10,.6287)
		};

		\addplot coordinates { %[mark=square] coordinates {
			(0, .6163)	
			(1, .6163)
			(2, .6163)
			(3, .6109)
			(4, .599)
			(5, .5812)
			(6, .5567)
			(7, .5266)
			(8, .491)
			(9, .4508)
			(10, .4064)
		};

		\addplot coordinates { %[mark=square] coordinates {
			(0,.5859)
			(1,.5859)
			(2,.5859)
			(3,.5341)
			(4,.4696)
			(5,.3961)
			(6,.3188)
			(7,.2401)
			(8,.1607)
			(9,.08088)
			(10,.004861)
		};

		\addplot coordinates { %[mark=star] coordinates {
			(0,.3787)
			(1,.3787)
			(2,.3787)
			(3,.3192)
			(4,.2671)
			(5,.2296)
			(6,.2152)
			(7,.2152)
			(8,.2152)
			(9,.2152)
			(10,.2152)
		};

		\addplot coordinates { %[mark=o] coordinates {
			(0,.7538)
			(1,.7538)
			(2,.7538)
			(3,.6885)
			(4,.6231)
			(5,.5578)
			(6,.4925)
			(7,.4925)
			(8,.4925)
			(9,.4925)
			(10,.4925)
		};

   		\addplot coordinates { %[mark=square*] coordinates {
   			(0,.5115)
   			(1,.5115)
   			(2,.5115)
   			(3,.4477)
   			(4,.3838)
   			(5,.3200)
   			(6,.2561)
   			(7,.2561)
   			(8,.2561)
   			(9,.2561)
   			(10,.2561)
   		};

@@

@@

	\section{Guidelines: Using the Parallel Spatial Locality Metric}

	-- Notes

	 GEM: The recorded parallel spatial locality approaches the theoretical limit of 0 (0.0124 at 10 bits dropped). This indicates that almost all memory accesses made by the kernel are perfectly synchronized between OpenCL threads.

	 LUD: Codes with a mix of favourable and unfavourable access patterns (typically both dimensions of matrices) will have a steep dropoff to ~0.5 of the starting PSL metric, and flattens out at that point.

	%TODO: Compile list of guidelines

	\begin{comment}
	\section{Predicting OpenCL Kernel Performance using AIWC}

	\begin{comment}
	- AIWC performance prediction paper
	- Important for workload schedulers and compiler writers.
	- Used in this paper to demonstrate PSL is inherently linked to performance-prediction.

	-- Methodology:
		- Leave-one-out predictive accuracy
		- CLgen?
			- intro
			- use to generate kernels for performance prediction

	In~\cite{beauperformanceprediction} we show that AIWC features can be used to build performance prediction models for the runtime of OpenCL codes on diverse heterogeneous platforms.
	This is useful for workload schedulers at supercomputer facilities and compiler engineers who must predict performance of codes on architectures without gathering performance data directly.
	In this paper, we test whether the implemented parallel spatial locality metric is linked to code performance by building a predictive model

	- Used random forest model (ranger).

	- 
	\end{comment}

@@

	To the best of our knowledge, this work is the first to propose the use of architecture-independent metrics of parallel memory access to characterise hardware-specific optimization. 
	We implemented and evaluated a new feature, relative local memory usage, to help characterize memory-based GPU code optimizations.
	We also implemented a new parallel spatial locality metric to capture the idea of closeness of memory accesses made by parallel OpenCL workloads.
	We ran the enhanced AIWC tool on matrix multiply kernels and selected OpenDwarfs benchmarks, presenting results and analysis to validate our methodology. 
	%Our work is the first to measure memory access patterns of parallel codes in an architecture-independent, easily interpreted manner. %Josh, Beau: is this too bold a claim? If unnecessary, please remove.
	%TODO UPDATE IN FUTURE

@@

	While our work in this paper intended to guide optimization efforts, another potential use case for AIWC is providing architecture-independent performance predictions for OpenCL kernels by generating machine learning models based on AIWC metrics \cite{beauperformanceprediction}.  Future work would modify the presented metrics and develop new memory metrics with the specific intent of being fed to machine learning models to predict kernel performance on arbitrary and novel architectures. %FIXME: I want to say that we have been focusing on _characterising optimization_. We can do more work in making the AIWC metrics better for _performance prediction_. Is this coming across okay in this paragraph?
